---
title: "PartiTargetPipeline"
author: "Susan Garcia"
date: "2023-10-04"
output: html_document
---
somethings wrong..... :(

Goals:

Section I: Import and cleaning/rearranging of data.
Section II: Applying Best-Matched Internal Standard (B-MIS).
Section III: Quantifying peak area to umol/vial when possible. 


Loading libraries
```{r}
library(tidyverse)
library(janitor)
library(dplyr)
library(tidyr)
library(readr)
```
# Section I: Import and cleaning/rearranging of data.
## Cleaning transition list
```{r}
transition_list <- read_csv("data_raw/HILIC_QE_Positive_Parti_CRR_TransitionResults.csv")

#Cleaning up the transition list
clean_areas <- transition_list %>%
  select(filename=`Replicate Name`, compound_name=`Precursor Ion Name`, area=Area) %>%
  mutate(cmpd_type=ifelse(str_detect(compound_name, ", "), "IS", "Non-IS")) %>%
  mutate(area=as.numeric(area)) %>%
  mutate(samp_type=str_extract(filename, "Poo|Blk|Smp|Std")) %>%
  mutate(day=str_extract(filename, "D\\d+")) %>%
  mutate(day=as.numeric(str_remove(day, "D"))) %>%
  arrange(day) %>%
  mutate(filename=fct_inorder(filename)) %>%
  select(-day) %>%
  arrange(compound_name, filename)

#Making a df for just the IS
IS_list <- clean_areas %>%
  filter(cmpd_type=="IS")
```
# Importing Standards List
```{r}
standards_list <- read_csv("data_raw/Ingalls_Lab_Standards.csv") %>%
  clean_names() %>%
  select("compound_name", "hilic_mix", "concentration_u_m", "z") %>%
  filter(z == 1) %>%
  mutate(cmpd_type=ifelse(str_detect(compound_name, ", "), "IS", "Non-IS")) %>%
  semi_join(clean_areas, by = "compound_name") %>%
  filter (cmpd_type == "Non-IS") %>%
  rename(mix = hilic_mix, conc_um = concentration_u_m)
```


# Section II: Applying Best-Matched Internal Standard (B-MIS).
```{r}
#Performing a visual check on the IS
clean_areas %>%
  filter(compound_name=="Homarine" |
           cmpd_type=="IS") %>%
  filter(samp_type=="Poo") %>%
  group_by(compound_name) %>%
  ggplot() +
  geom_col(aes(x=filename, y=area)) +
  facet_wrap(~compound_name, ncol=2, scales="free_y")

# Not sure yet
chosen_BMIS <- clean_areas %>%
  filter(cmpd_type=="Non-IS") %>%
  filter(samp_type=="Poo") %>%
  # summarise(cv_init=sd(area)/mean(area))
  left_join(IS_list, by="filename", suffix=c("", "_IS"), relationship = "many-to-many") %>%
  select(filename, compound_name, compound_name_IS, area, area_IS) %>%
  group_by(compound_name, compound_name_IS) %>%
  mutate(norm_area=area/area_IS*mean(area_IS)) %>%
  summarise(cv_IS=sd(norm_area)/mean(norm_area)) %>%
  arrange(compound_name, cv_IS) %>%
  slice(1)


#Not sure yet
BMISed_areas <- clean_areas %>%
  filter(cmpd_type=="Non-IS") %>%
  select(filename, compound_name, area) %>%
  left_join(chosen_BMIS) %>%
  select(-cv_IS) %>%
  left_join(clean_areas, by=c("filename", compound_name_IS="compound_name"),
            suffix=c("", "_IS")) %>%
  # slice(1:43) %>% distinct(filename) %>% print(n=Inf)
  select(filename, compound_name, area, area_IS) %>%
  group_by(compound_name) %>%
  mutate(bmised_area=area/area_IS*mean(area_IS[1:39], na.rm = TRUE)) %>%
  select(filename, compound_name, bmised_area) 
```


# Section III: Quantifying peak area to umol/vial when possible.

```{r}
all_rfs <- clean_areas %>%
  left_join(standards_list, by = "compound_name") %>%
  #mutate(bmised_area = as.numeric(area)) %>%
  filter(str_detect(filename, paste0(mix, "InMatrix|H2OinMatrix"))) %>%
  mutate(std_type = str_extract(filename, "Mix|H2O")) %>%
  select(-mix) %>%
  group_by(compound_name, std_type, conc_um) %>%
  summarise(mean_area = mean(area)) %>%
  pivot_wider(names_from = std_type, values_from = mean_area) %>%
  mutate(area_from_4um_addition = Mix - H2O) %>%
  mutate(rf = conc_um / area_from_4um_addition) %>%
  mutate(rf_ratio = Mix / H2O) %>%
  select(compound_name, rf, rf_ratio)

```

```{r}
parti_final_concs <- BMISed_areas %>%
  filter(!str_detect(filename, "Std")) %>%
  left_join(all_rfs) %>%
  mutate(uM_in_vial=bmised_area/rf) %>%
  select(compound_name, filename, uM_in_vial) %>%
  mutate(uM_in_env=uM_in_vial/100) %>%
  mutate(nM=uM_in_env*1000) %>%
  select(compound_name, filename, nM)
```
## Write out the CSV for all
```{r}
write_csv(parti_final_concs, "intermediates/full_parti_final_concs.csv")
```

